\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-01-14 honors_thesis/thesis/paper.tex

\documentclass[conference,letterpaper]{IEEEtran}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{psfrag}

\IEEEoverridecommandlockouts

@@

\title{\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization%
\thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department
of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602
(phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}%
}

\date{}

\author{Matthew Gardner, Andrew McNabb, and Kevin Seppi}

\begin{document}
\maketitle

\begin{abstract}

Particle swarm optimization (PSO) has previously been parallelized only by 
adding more particles to the swarm or by parallelizing the evaluation of the
objective function.  However, some functions are better optimized with more
iterations and fewer particles.  Accordingly, we take inspiration from 
speculative execution commonly performed in modern processors and propose what
we call speculative evaluation in PSO.  Future positions of the particles are
speculated and evaluated in parallel with current positions, performing two
iterations of PSO at once.

We also propose another way of making use of these speculative particles,
keeping the best position found instead of the position that PSO actually would
have taken.  We show results that for a number of functions, speculative
evaluation can have dramatic improvements over adding additional particles to
the swarm.

\end{abstract}

@@

There are some functions for which almost no exploration needs to be done;
Sphere is an example of such a function, as would be a function where the
general area of the optimal point is known, but the exact point needs to be
found.  For such functions the best use of processors is to have a small swarm
performing speculative evaluation with our second method, where all speculative
evaluations are used.

>>>>

\documentclass[oneside,honors]{honors}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{psfrag}
\pagestyle{plain}

@@

\TitleTop{Speculative Evaluation in}
\TitleBottom{Particle Swarm Optimization}
\Year{2010}
\Month{April}
\degree{Bachelor of Science}

\Author{Matthew Gardner}
\Advisor{Kevin Seppi}
\Abstract{
Particle swarm optimization (PSO) has previously been parallelized only by 
adding more particles to the swarm or by parallelizing the evaluation of the
objective function.  However, some functions are better optimized with more
iterations and fewer particles.  Accordingly, we take inspiration from 
speculative execution commonly performed in modern processors and propose what
we call speculative evaluation in PSO.  Future positions of the particles are
speculated and evaluated in parallel with current positions, performing two
iterations of PSO at once.

We also propose another way of making use of these speculative particles,
keeping the best position found instead of the position that PSO actually would
have taken.  We show that for a number of functions, speculative evaluation
gives dramatic improvements over adding additional particles to the swarm.
}

\begin{document}
\frontmatter
\makepreliminarypages

\doublespace
\tableofcontents
\clearemptydoublepage

\singlespace
\listoffigures
\clearemptydoublepage

\doublespace

\mainmatter

@@

There are some functions for which very little exploration needs to be done;
Sphere is an example of such a function, as would be a function where the
general area of the optimal point is known, but the exact point needs to be
found.  For such functions the best use of processors is to have a small swarm
performing speculative evaluation with our second method, where all speculative
evaluations are used.
